---
title: "Inequitable Access"
subtitle: "An Analysis of Licenced Child Care in Toronto's 25 Wards"
author: 
  - Thomas Fox
thanks: "Code and data are available at: https://github.com/ThomasWilliamFox/child_care_access.git."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
toc: True 
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(arrow)
library(here)
library(palmerpenguins)
library(ggmap)
library(maps)
library(knitr)
```

```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false

#### Load data ####

# read child care data

child_care_data = read_parquet(
  file = here("data/analysis_data/child_care_data.parquet"),
  show_col_types = FALSE
)

# read summarized child care data 

summarized_child_care_data = read_parquet(
  file = here("data/analysis_data/ward_child_care_data.parquet"),
  show_col_types = FALSE
)

# read census ward data 

ward_data = read_parquet(
  file = here("data/analysis_data/census_data.parquet"),
  show_col_types = FALSE
)

# read merged data 

merged_ward_childcare = read_parquet(
  file = here("data/analysis_data/merged_ward_data.parquet"),
  show_col_types = FALSE
)

# read names data

ward_names = read_parquet(
  file = here("data/analysis_data/ward_names.parquet"),
  show_col_types = FALSE
)
```

# Introduction

Using @citeR and @rohan.

The paper's estimand is that average household income has a negative relationship with the number of children per child care space by ward. 

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}
## Licensed Child Care Centres {#sec-childcare}

```{r}
#| label: tbl-childcaredata
#| tbl-cap: Sample of Cleaned Toronto Licensed Child Care Centre Data
#| echo: false
#| warning: false

child_care_table <- 
  child_care_data |>
  select(x_id, ward, totspace, auspice, subsidy, cwelcc_flag)

head(child_care_table, 6) |>
  kable(
    col.names = c("Facility ID", " Ward Number", "Total Spaces", "Type", "Subsidy", "SWELCC"),
    booktabs = TRUE
  )

```

```{r out.width = '100%'}
#| label: fig-map
#| fig-cap: Child care centres in Toronto
#| warning: false
#| message: false
#| echo: false

# Use google maps for box co-ordinates
bbox_toronto <- c(left =-79.669692514001, bottom = 43.57044, right = -79.096000099990, top = 43.85090373064596)

# Adds API key following instructions found here: https://tellingstorieswithdata.com/05-static_communication.html#maps
register_stadiamaps(key = "1662b375-446a-4847-acf3-831fb1381f2b", write = TRUE)

# Builds map using box co-ordinates
toronto_stamen_map <- get_stadiamap(bbox_toronto, zoom = 12, maptype = "stamen_toner_lite", crop = TRUE)

# Builds map with child care centres plotted
ggmap(toronto_stamen_map, extent = "panel", maprange = FALSE) +
  geom_point(data = child_care_data,
             aes(x = as.double(x), y = as.double(y), colour = ward),
             alpha = 0.7, size = log(child_care_data$totspace)*.22, show.legend = FALSE) +
  labs(x = "Longitude", y = "Latitude") + 
  theme(axis.text.x = element_text(size = 8),
        axis.text.y = element_text(size = 8))
  
```

## Toronto Ward Profiles {#sec-warddata}

```{r}
#| label: tbl-warddata
#| tbl-cap: Sample of Cleaned Toronto Ward Data
#| echo: false
#| warning: false

ward_data_table <- 
  merged_ward_childcare |>
  select(ward, total, total_under_15, avg_hh_income, med_hh_income, english, visible_minority)

head(ward_data_table, 6) |>
  kable(
    col.names = c("Ward", "Total Population", "Children under 15", "Average Household Income", "Median Household Income", "English spoken most often in household", "Population identifying as racialized"),
    booktabs = TRUE
  )

```

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.  

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a negative relationship between average household income and the number of children per child care space by ward. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

```{r}
#| label: fig-incomespaces
#| fig-cap: Relationship between income and child care spaces
#| warning: false
#| message: false
#| echo: false
merged_ward_childcare |>
  ggplot(aes(x = avg_hh_income, y = total_under_15/total_spots)) +
  geom_point(fill = "black", size = 1, shape = 23) +
  geom_smooth(method = lm, color = "black") +
  theme_minimal() +
  labs(x = "Average income by ward", y = "Children per child care space") 
```

```{r}
#| label: fig-languagespaces
#| fig-cap: Relationship between language and child care spaces
#| warning: false
#| message: false
#| echo: false
merged_ward_childcare |>
  ggplot(aes(x = english/total*100, y = total_under_15/total_spots)) +
  geom_point(fill = "black", size = 1, shape = 23) +
  geom_smooth(method = lm, color = "black") +
  theme_minimal() +
  labs(x = "Proportion of English speaking households by ward (%)", y = "Children per child care space") 
```

```{r}
#| label: fig-nonminorityspaces
#| fig-cap: Relationship between non-racialized population by ward and child care spaces
#| warning: false
#| message: false
#| echo: false
merged_ward_childcare |>
  ggplot(aes(x = ((total-visible_minority)/total)*100, y = total_under_15/total_spots)) +
  geom_point(fill = "black", size = 1, shape = 23) +
  geom_smooth(method = lm, color = "black") +
  theme_minimal() +
  labs(x = "Porportion of ward population identified as non-racialized (%)", y = "Children per child care space") 
```  

Our results are summarized in @tbl-modelresults.

```{r}
#| include: false
#| echo: false
#| eval: true
#| warning: false
#| message: false

#library(rstanarm)

#first_model <-
#  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| include: false
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

#modelsummary::modelsummary(
#  list(
#    "First model" = first_model
#  ),
#  statistic = "mad",
#  fmt = 2
#)
```


# Discussion

## First discussion point {#sec-first-point}

The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps


\newpage

\appendix

# Appendix {-}


# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows... 

```{r}
#| include: false
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

#pp_check(first_model) +
#  theme_classic() +
#  theme(legend.position = "bottom")

#posterior_vs_prior(first_model) +
#  theme_minimal() +
#  scale_color_brewer(palette = "Set1") +
#  theme(legend.position = "bottom") +
#  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| include: false
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

#plot(first_model, "trace")

#plot(first_model, "rhat")
```



\newpage


# References


