---
title: "Inequitable Access"
subtitle: "An Analysis of Licensed Child Care in Toronto's 25 Wards"
author: 
  - Thomas Fox
thanks: "Code and data are available at: https://github.com/ThomasWilliamFox/child_care_access.git."
date: today
date-format: long
abstract: "Child care is essential for the well being of families, communities, and the children able to attend these facilities. This paper examines the accessibility of licensed child care centres across Toronto's 25 wards. Findings indicate that there are more children per existing child care space in wards with lower household incomes, lower proportions of English speaking households, and higher proprotions of the population identifying as racialized. These findings suggest inequitable access to licensed child care spaces based on a variety of social factors, supporting initiatives aimed at increasing access to licensed child care in the city of Toronto."
format: pdf
number-sections: true
bibliography: references.bib
toc: True 
nocite: |
  @tellingstories
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(arrow)
library(here)
library(palmerpenguins)
library(ggmap)
library(maps)
library(knitr)
library(rstanarm)
```

```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false

#### Load data ####

# read child care data

child_care_data = read_parquet(
  file = here("data/analysis_data/child_care_data.parquet"),
  show_col_types = FALSE
)

# read summarized child care data 

summarized_child_care_data = read_parquet(
  file = here("data/analysis_data/ward_child_care_data.parquet"),
  show_col_types = FALSE
)

# read census ward data 

ward_data = read_parquet(
  file = here("data/analysis_data/census_data.parquet"),
  show_col_types = FALSE
)

# read merged data 

merged_ward_childcare = read_parquet(
  file = here("data/analysis_data/merged_ward_data.parquet"),
  show_col_types = FALSE
)


merged_data <- read_parquet(here::here("data/analysis_data/merged_ward_data.parquet"))
merged_data <- merged_data |>
  mutate(income = avg_hh_income / max(avg_hh_income))

# read names data

ward_names = read_parquet(
  file = here("data/analysis_data/ward_names.parquet"),
  show_col_types = FALSE
)
```

# Introduction

Equitable access to child care is essential to the social and economic health of a community such as the City of Toronto. Access to child care has been shown have a positive impact occupational and educational opportunities for parents, especially those in lower income brackets [@childcare]. These economic advantages for parents and families bring benefits to their physical and social well-being [@childcare]. Child care access also impacts the health and development of children attending these facilities [@rhijn]. Vulnerable children disproportionately benefit from child care access as these centers help to facilitate early intervention methods [@underwood]. Equitable access to child care is therefore a vital facet of community health and development.

This paper explores child care data from @caredata and ward data from @wards to explore child care access across the city of Toronto. These data sets are made freely available by Open Data Toronto [@opendatatoronto]. Data analysis and processing was performed using the programming language `R` [@citeR]. Various packages were used throughout the writing of this paper including `tidyverse` [@tidyverse], `arrow` [@arrow], `here` [@here], `ggmap` [@ggmap], `maps` [@maps], `knitr` [@knitr], `dplyr` [@dplyr], `janitor` [@janitor], and `rstanarm` [@rstanarm].    

This analysis explores relationships between child care demand and various demographic metrics across Toronto's 25 wards (see [@sec-results]). The paper's estimand is that average household income is negatively correlated with the number of children per child care space by ward. The paper also explores the relationships between existing child care spaces and the proportion of racialized and non-native-English speaking populations by ward. These findings demonstrate inequitable access to child care across the city of Toronto. As child care plays a central role in the social and economic well-being of communities, and has an especially positive impact on vulnerable children and low-income families, these findings support measures and initiatives aimed at ensuring more equitable access to child care in the city of Toronto [see @strategy].

The remainder of this paper is structured as follows. @sec-data outlines the data sets used throughout this paper made available by Open Data Toronto [@opendatatoronto]. @sec-childcare explores the "Licensed Child Care Centres" data set provided by @caredata, while @sec-warddata explores the "Ward Profiles (25-Ward Model)" data set provided by @wards. A model built to explore the impact of average household income on child care spaces in Toronto is outlined in @sec-model. Results of data analysis are found in @sec-results. This is followed by a discussion surrounding these findings and their implications in @sec-discussion. This section also includes discussions around potential weaknesses and next steps regarding this research.  

# Data {#sec-data}
## Licensed Child Care Centres {#sec-childcare}

The "Licensed Child Care Centres" data set is provided to Open Data Toronto [@opendatatoronto] by the City of Toronto's Children's Services division [@caredata]. The data is updated on an ongoing basis and the data used in this paper was updated on April 19, 2024. The data set contains 1,066 entries, each corresponding to a licensed child care facility in Toronto. Variables chosen to isolate after cleaning include the facility ID number, the ward number where the facility is located, and the total number of individual child care spaces at the facility. Each facility's operation type is also isolated, with the data showing that there are 703 non profit, 324 commercial, and 39 public (City operated) facilities in Toronto. Information about whether the each facility has a fee subsidy contract or participates in the Canada-Wide Early Learning & Child Care (CWELCC) system is also included, with 68% of facilities having subsidy contracts and 87% of facilities participating in CWELCC. @tbl-childcaredata shows the first six entries in this data set and @fig-map shows the location of each facility across Toronto. 

```{r}
#| label: tbl-childcaredata
#| tbl-cap: Sample of Cleaned Toronto Licensed Child Care Centre Data
#| echo: false
#| warning: false
#| fig.pos: "H"

child_care_table <- 
  child_care_data |>
  select(x_id, ward, totspace, auspice, subsidy, cwelcc_flag)

head(child_care_table, 6) |>
  kable(
    col.names = c("Facility ID", " Ward Number", "Total Spaces", "Type", "Subsidy", "CWELCC"),
    booktabs = TRUE,
    align = "l"
  )

```

```{r out.width = '100%'}
#| label: fig-map
#| fig-cap: Map showing the loction of all licensed child care centres in Toronto
#| warning: false
#| message: false
#| echo: false
#| fig.pos: "H"

# Use google maps for box co-ordinates
bbox_toronto <- c(left =-79.669692514001, bottom = 43.57044, right = -79.096000099990, top = 43.85090373064596)

# Adds API key following instructions found here: https://tellingstorieswithdata.com/05-static_communication.html#maps
register_stadiamaps(key = "1662b375-446a-4847-acf3-831fb1381f2b", write = TRUE)

# Builds map using box co-ordinates
toronto_stamen_map <- get_stadiamap(bbox_toronto, zoom = 12, maptype = "stamen_toner_lite", crop = TRUE)

# Builds map with child care centres plotted
ggmap(toronto_stamen_map, extent = "panel", maprange = FALSE) +
  geom_point(data = child_care_data,
             aes(x = as.double(x), y = as.double(y)),
             alpha = 0.7, size = log(child_care_data$totspace)*.20, show.legend = FALSE) +
  labs(x = "Longitude", y = "Latitude") + 
  theme(axis.text.x = element_text(size = 8),
        axis.text.y = element_text(size = 8))
  
```

@fig-totalspots displays the total individual child care spaces found in each of Toronto's 25 wards. The wards with the fewest number of licensed child care spaces are Scarborough-Rouge Park with 1935 spaces, Etobicoke-North with 2094, and Scarborough North with 2104. The wards with the greatest number of child care spaces are Parkdale-High Park with 4632, Etobicoke-Lakeshore with 4937, and Toronto-Danforth with 5054.

```{r}
#| label: fig-totalspots
#| fig-cap: Total Number of Licensed Child-Care Spaces in Toronto by Ward
#| echo: false
#| warning: false
#| fig.pos: "H"

# turn ward_names into list for graph substitution with ward numbers 
ward_name_list <- as.list(ward_names$ward_name)

# substitute ward numbers for names
ward_childcare <-
  merged_ward_childcare |>
  mutate(ward = as.character(ward))

for (x in 1:25) {
  ward_childcare$ward[ward_childcare$ward == x] = ward_name_list[[x]]
} 

# construct graph 
ward_childcare |>
  ggplot(aes(x = reorder(ward, 1:25), y = total_spots)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 7, angle = 45, vjust = 1, hjust=1))+
  labs(x = "Ward", y = "Total child care spaces")

```

## Toronto Ward Profiles {#sec-warddata}

The "Ward Profiles (25-Ward Model)" data set is provided to Open Data Toronto (@opendatatoronto) by Toronto City Planning [@wards]. The data sets of interest in this paper found through this resource are the "2023-WardProfiles-2011-2021-CensusData" data set which is used to determine demographic information related to each ward, and the "25-WardNames-Numbers" data set that is used to code names to ward numbers in this paper [@wards]. Variables isolated from the 2021 Canada Census include ward number, population, number of children under 15, average yearly household income, number of households where English is spoken most often, population identifying as racialized. @tbl-warddata shows the first six entries in the cleaned data set.

```{r}
#| label: tbl-warddata
#| tbl-cap: Sample of the First Six Entries in the Cleaned Toronto Ward Data
#| echo: false
#| warning: false
#| fig.pos: "H"

ward_data_table <- 
  merged_ward_childcare |>
  select(ward, total, total_under_15, avg_hh_income, english, visible_minority)

head(ward_data_table, 6) |>
  kable(
    col.names = c("Ward", "Total Population", "Children Under 15", "Average Household Income", "English Spoken Most Often in Household", "Population Identifying as Racialized"),
    booktabs = TRUE,
    align = "l"
  )

```

@fig-totalchildren displays the total number of children under the age of 15 in each ward. Age range was selected to represent the child care age ranges found on the Government of Ontario's child care rules website [@ontario]. The wards with the fewest number of children are University-Rosedale with 8980, Spadina-Fort York with 9270, and Toronto Centre with 9310. The wards with the greatest number of children are Beaches-East York with 18685, Humber River-Black Creek with 18770, and Eglinton-Lawrence with 19115.

```{r}
#| label: fig-totalchildren
#| fig-cap: Population Under 15 Years of Age in Each of Toronto's 25 Wards
#| echo: false
#| warning: false
#| fig.pos: "H"

# substitute ward numbers for names
for (x in 1:25) {
  merged_ward_childcare$ward[merged_ward_childcare$ward == x] = ward_name_list[[x]]
} 

# construct graph 
merged_ward_childcare|>
  ggplot(aes(x = reorder(ward, 1:25), y = total_under_15)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 7, angle = 45, vjust = 1, hjust=1))+
  labs(x = "Ward", y = "Population under 15 years of age") 
```

@fig-avgincome shows the average yearly household income in each of Toronto's 25 wards. The wards with the lowest average income in Toronto are Humber River-Black Creek with \$85700, York South-Weston with \$88700, and Toronto Centre with \$89400. Toronto's wards with the highest average income are University-Rosedale with \$174800, Eglinton-Lawrence with \$176400, and Don Valley West with \$224800.

```{r}
#| label: fig-avgincome
#| fig-cap: Average Household Income in Each of Toronto's 25 Wards
#| echo: false
#| warning: false
#| fig.pos: "H"

# Construct graph 
merged_ward_childcare |>
  ggplot(aes(x = reorder(ward, 1:25), y = avg_hh_income)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 7, angle = 45, vjust = 1, hjust=1)) +
  labs(x = "Ward", y = "Average household income")
```
```{r}
#| label: fig-language
#| fig-cap: Proportion of the Population in Each of Toronto's 25 Wards That Speak English Most Often at Home
#| echo: false
#| warning: false
#| fig.pos: "H"

# Construct graph 
merged_ward_childcare |>
  ggplot(aes(x = reorder(ward, 1:25), y = english/total*100)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 7, angle = 45, vjust = 1, hjust=1)) +
  labs(x = "Ward", y = "English speaking households (%)")
```

@fig-language shows the proportion of the population in each of Toronto's 25 wards that speaks English most often in their households. The wards with the lowest proportion of English speaking households in Toronto are Humber River-Black Creek with 41%, York South-Weston with 43%, and Toronto Centre with 48%. Toronto's wards with the highest proportion of English speaking households are Toronto-Danforth with 81%, Beaches-East York with 81%, and Park Dale-High Park with 82%.

```{r}
#| label: fig-racialized
#| fig-cap: Proportion of the Population in Each of Toronto's 25 Wards Identify as Racialized
#| echo: false
#| warning: false
#| fig.pos: "H"

# Construct graph 
merged_ward_childcare |>
  ggplot(aes(x = reorder(ward, 1:25), y = visible_minority/total*100)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 7, angle = 45, vjust = 1, hjust=1)) +
  labs(x = "Ward", y = "Racialized population (%)")
```

@fig-racialized shows the proportion of the population in each of Toronto's 25 wards that identifies as racialized. The wards with the lowest proportion of racialized population in Toronto are Parkdale-High Park with 29%, Etobicoke Centre with 32%, and Davenport with 34%. Toronto's wards with the highest proportion of racialized population are Etobicoke-North 78%, Scarborough-Agincourt with 82%, and Scarborough-North with 92%.

# Model {#sec-model}
```{r}
#| include: false
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

The goal of this paper's modelling strategy is to investigate the effects that average household income has on the number of children per existing child care space in each of Toronto's 25 wards.

Further model details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of children per child care space in a ward. Then $\beta_i$ is the average household income in the ward.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \beta_i \\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

The model is ran in R [@citeR] using the `rstanarm` package of @rstanarm with the package's default prior settings.

### Model justification

A negative relationship is expected between average household income and the number of children per child care space by ward. 

### Model Results 

The model results are summarized in @tbl-modelresults. The model shows that the intercept, or children per child care space if $\beta_i$ were 0, would be 8.12. The model also shows that the impact of average household income being 1.0, or the maximum average income in Toronto's 25 wards is -5.64 children per child care space. This infers that average income has a negative relationship with the number of children per existing child care space in the city of Toronto. 

```{r}
#| include: true
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false
#| fig.pos: "H"

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```


# Results {#sec-results}

To explore the effect that average household income, language, and racialized population have on the total number of child care spaces by ward, a merged and summarized data set was created with variables for ward name, average yearly household income, total childcare spaces, and total population aged 0 to 14, total population, population identifying as racialized, English speaking population, and children per child care space (see @tbl-mergeddata). @fig-childrenspaces displays the number of children per existing child care space by ward. These values help infer the accessibility and demand for child care spaces in each ward. The wards with the greatest number of children per existing child care space are Etobicoke North with 8.8, Scarborough-Rouge Park with 7.8, and York South-Weston with 7.7. The wards with the fewest number of children per child care space are Toronto-Danforth with 3.2, Spadina-Fort York with 2.9, and University-Rosedale with 2.7.

```{r}
#| label: fig-childrenspaces
#| fig-cap: Number of Children for Every Existing Licensed Child Care Space in Each of Toronto's 25 Wards
#| echo: false
#| warning: false
#| fig.pos: "H"

# Construct graph 
merged_ward_childcare |>
  ggplot(aes(x = reorder(ward, 1:25), y = prop)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 7, angle = 45, vjust = 1, hjust=1)) +
  labs(x = "Ward", y = "Children per child care space")
```

@fig-incomespaces aims to assess the relationship between average household income and children per child care space in Toronto's 25 wards. The plot displays a negative correlation between average household income and children per child care space. 

```{r}
#| label: fig-incomespaces
#| fig-cap: Relationship between income and child care spaces
#| warning: false
#| message: false
#| echo: false
#| fig.pos: "H"

merged_ward_childcare |>
  ggplot(aes(x = avg_hh_income, y = total_under_15/total_spots)) +
  geom_point(fill = "black", size = 1, shape = 23) +
  geom_smooth(method = lm, color = "black") +
  theme_minimal() +
  labs(x = "Average household income by ward", y = "Children per child care space") 
```

@fig-languagespaces explores the relationship between English speaking household populations and children per child care space in Toronto's 25 wards. The plot displays a negative correlation between the proportion of English speaking populations and children per child care space. 

```{r}
#| label: fig-languagespaces
#| fig-cap: Relationship between language and child care spaces
#| warning: false
#| message: false
#| echo: false
#| fig.pos: "H"

merged_ward_childcare |>
  ggplot(aes(x = english/total*100, y = total_under_15/total_spots)) +
  geom_point(fill = "black", size = 1, shape = 23) +
  geom_smooth(method = lm, color = "black") +
  theme_minimal() +
  labs(x = "Proportion of ward population that speak English at home (%)", y = "Children per child care space") 
```

@fig-nonracializedspaces aims to assess the relationship between non-racialized populations and children per child care space in Toronto's 25 wards. The plot shows a negative correlation between proportion populations identifying as non-racialized and children per child care space. 

```{r}
#| label: fig-nonracializedspaces
#| fig-cap: Relationship between non-racialized population by ward and child care spaces
#| warning: false
#| message: false
#| echo: false
#| fig.pos: "H"

merged_ward_childcare |>
  ggplot(aes(x = ((total-visible_minority)/total)*100, y = total_under_15/total_spots)) +
  geom_point(fill = "black", size = 1, shape = 23) +
  geom_smooth(method = lm, color = "black") +
  theme_minimal() +
  labs(x = "Porportion of ward population identified as non-racialized (%)", y = "Children per child care space") 
```  

# Discussion {#sec-discussion}



## Average Household Income



## English Speaking and Racialized Population



## Weaknesses and Next Steps


\newpage

\appendix

# Appendix {-}

```{r}
#| label: tbl-mergeddata
#| tbl-cap: Sample of Merged Licensed Child Care Centres in Toronto and Ward Profiles (2021 Census Data) used to build graphs
#| echo: false
#| warning: false
#| layout-nrow: 2

merged <-
  merged_ward_childcare |>
  select(ward, avg_hh_income, total_spots, total_under_15, total, visible_minority, english, prop)

head(merged[c(1:4)], 3) |>
  kable(
    col.names = c("Ward Name", "Average Household Income", "Child Care Spaces", "Population Aged 0 - 14"),
    booktabs = TRUE,
    align = "l"
  )

head(merged[c(5:8)], 3) |>
  kable(
    col.names = c("Total Population", "Racialized Population", "English Speakng Population", "Children Per Space"),
    booktabs = TRUE,
    align = "l"
  )

```

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows that the model can be used to simulate data similar to our source data.

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows how much the model's estimates change once the source data is incorporated. 

```{r}
#| include: true
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-plot displays the 95% credibility interval for the intercept, average income, and standard deviation (sigma)

@fig-stanareyouokay-1 is a trace plot showing that after checking the model for inconsistencies, the model appears to be working properly.

@fig-stanareyouokay-2 is a Rhat plot, similarly showing that the model appears to be working properly.

```{r}
#| include: true
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-plot
#| fig-cap: "Checking the credibility of the model"

plot(
  first_model,
  "areas"
)

```


```{r}
#| include: true
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```



\newpage


# References


